### 2017-10-17 Cars/basicModel.Rnw
general updates

\begin{figure}[hb]
\centering
\includegraphics[width = 0.6\textwidth]{basicFit.png}
\end{figure}

@@

The complexity of the steering process in (\ref{dEq}) is increased by introducing a two-state Markov switching variable $s_t$, where $s_t = 0$ indicates that the driver is continuing on their current path, and $s_t = 1$ will turn the car towards the centre of the lane. The new steering equation is described by
\begin{equation}
\delta_t = (1 - s_t) \delta_{t-1}  s_t (pi/2 + \zeta(x_{t-1} - M)) +  \eta_t \label{dEqMS}.
\end{equation}
where $M$ denotes the midpoint of the lane. Note that this now describes a random walk when $s_t = 0$, with large departures from $x_t = M$ controlled when $s_t = 1$.  $s_t$ is governed by the transition probabilities $\rho_{ij} = p(s_t = j | s_{t-1} = 1)$, with $\rho_{00} = 1 - \rho_{01}$ and $\rho_{10} = 1 - \rho_{11}$. As drivers are more likely to steer towards the centre of the lane the further away they are from the centre, $\rho_{10}$ and $\rho_{11}$ are given by 
\begin{equation}
\rho_{i1} = \frac{1}{1 + \exp\{-(\alpha_i + \beta_i (x_{t-1} - M)^2)\}} \mbox{ for } i = 0, 1.
\end{equation}
With this formulation, $s_t =  0$ represents a default driving state, and $s_t = 1$ represents the driver actively correcting for drift.

This formulation removes $\gamma$ from $\theta$, but adds five additional parameters, $\zeta, \alpha_0, \beta_0, \alpha_1$, and $\beta_1$, each with a $N(\mu_i, \tau_i^2)$ prior. The likelihood can be calculated by the Hamilton filter, and is given by
\begin{equation}
p(x_{4:T}, y_{4:T} | \theta) = \prod_{t=4}^T \left(\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta) \right)
\end{equation}
where
\begin{align}
p_{\delta 0}(x_t, y_t | \theta) &\propto  \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \delta(\Delta x_{t-1}, \Delta y_{t-1}) \right)^2 \right\} \label{dDensS0}, \\
p_{\delta 1}(x_t, y_t | \theta) &\propto   \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \pi/2 - \zeta (x_{t-1} - M)\right)^2  \right\} \label{dDensS1},
\end{align}
and
\begin{align}
\xi_{j, t} &= P(s_t = j | x_0, y_0, \dots, x_{t}, y_{t}, \theta) \\
&= \frac{\sum_{i=0}^1\rho_{ij} \xi_{i, t-1} p_v(x_{t}, y_{t})p_{\delta j}(x_{t}, y_{t})}
{\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta)} \label{HF}.
\end{align}
The recursion in (\ref{HF}) is initialised at $\xi_3 = \{0.9, 0.1\}$.
\end{document}

>>>>

\begin{figure}[hb]
\centering
\includegraphics[width = 1\textwidth]{basicFit.png}
\end{figure}

@@

The complexity of the steering process in (\ref{dEq}) is increased by introducing a two-state Markov switching variable $s_t$, where $s_t = 0$ indicates that the driver is continuing on their current path, and $s_t = 1$ will turn the car towards the centre of the lane. The new steering equation is described by
\begin{equation}
\delta_t = (1 - s_t) \delta_{t-1}  s_t (pi/2 + \zeta(x_{t-1} - M)) +  \eta_t \label{dEqMS}.
\end{equation}
where $M$ denotes the midpoint of the lane. Note that this now describes a random walk when $s_t = 0$, with large departures from $x_t = M$ controlled when $s_t = 1$.  $s_t$ is governed by the transition probabilities $\rho_{ij} = p(s_t = j | s_{t-1} = 1)$, with $\rho_{00} = 1 - \rho_{01}$ and $\rho_{10} = 1 - \rho_{11}$. As drivers are more likely to steer towards the centre of the lane the further away they are from the centre, $\rho_{10}$ and $\rho_{11}$ are given by 
\begin{equation}
\rho_{i1} = \frac{1}{1 + \exp\{-(\alpha_i + \beta_i (x_{t-1} - M)^2)\}} \mbox{ for } i = 0, 1.
\end{equation}
The velocity equation is also simplified for computational purposes (though the results had the same problems with it the same as earlier - so this did not really help).
\begin{equation}
v_t = v_{t-1} + \epsilon_t \label{vEqMS}
\end{equation}
With this formulation, $s_t =  0$ represents a default driving state, and $s_t = 1$ represents the driver actively correcting for drift.

This formulation removes $\gamma$ and $\phi$ from $\theta$, but adds five additional parameters, $\zeta, \alpha_0, \beta_0, \alpha_1$, and $\beta_1$, each with a $N(\mu_i, \tau_i^2)$ prior. The likelihood can be calculated by the Hamilton filter, and is given by
\begin{equation}
p(x_{4:T}, y_{4:T} | \theta) = \prod_{t=4}^T \left(\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta) \right)
\end{equation}
where
\begin{align}
p_v(x_t, y_t) &\propto \sigma_{\epsilon}^{-1} \exp\left\{ \frac{-1}{2 \sigma^2_{\epsilon}} \left(v(\Delta x_t, \Delta y_t) - v(\Delta x_{t-1}, \Delta y_{t-1})\right)^2\right\} \label{vDensS} \\
p_{\delta 0}(x_t, y_t | \theta) &\propto  \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \delta(\Delta x_{t-1}, \Delta y_{t-1}) \right)^2 \right\} \label{dDensS0}, \\
p_{\delta 1}(x_t, y_t | \theta) &\propto   \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \pi/2 - \zeta (x_{t-1} - M)\right)^2  \right\} \label{dDensS1},
\end{align}
and
\begin{align}
\xi_{j, t} &= P(s_t = j | x_0, y_0, \dots, x_{t}, y_{t}, \theta) \\
&= \frac{\sum_{i=0}^1\rho_{ij} \xi_{i, t-1} p_v(x_{t}, y_{t})p_{\delta j}(x_{t}, y_{t})}
{\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta)} \label{HF}.
\end{align}
The recursion in (\ref{HF}) is initialised at $\xi_2 = \{0.9, 0.1\}$.

\begin{figure}[hb]
\centering
\includegraphics[width = 1\textwidth]{VBMCMC}
\end{figure}
\end{document}
